---
title: "Appendix"
author: "Federico Chung"
date: "3/12/2021"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(tibble)
library(MASS) 
library(Metrics)
library(tidyr)
library(caret)
```

#Simulation Procedure

## Specify the TRUE Beta Values
We create a predetermined set of $\beta$:

```{r, eval=FALSE}
set.seed(2)
betas<- (sample(-10:10,size=50,replace=TRUE))/10
```

## CREATE THE SIMULATION VALUES

We then use the MVR Norm function which allows us to create X variables with a specified Covarience Matrix.

This allows our X-values to be

1. Normally Distributed explanatory variables
2. Multicollinear between the explanatory variables

The reason for the use of this simulation is that we wanted to simulate real world data where we tend to experience mutlicollinearity between the explanatory variables. Also we wanted to see if including multicollinearity our coefficient estimates for ridge, lasso, or elastic net can become an improvement to the estimates of the OLS model. 

```{r, eval = FALSE}
simulation<- function(seed, betas){
  set.seed(seed)
  n <- 100    # Number of observations
  p <- 50     # Number of predictors included in model
  CovMatrix <- outer(1:p, 1:p, function(x,y) {.7^abs(x-y)})
  x <- mvrnorm(n, rep(0,p), CovMatrix)
  
  e <- rnorm(n, 0, sd = 1)

  y <- x %*% betas + e
  df<- data.frame(y,x)
  return(df)
}
```

As it can be seen by the code chunk above, `CovMatrix` is where we specify the covariance matrix where we show that there is multicollinearity between the explanatory variables. 


## CREATING MODEL FUNCTIONS
We create getters for both the model, coefficient, and tuning parameters models for the following models:

- `OLS`

```{r, eval=FALSE}
## OLS
get_ols_model<- function(df){
  set.seed(45)
  ols_model<-train(
    y~.-1,
    data=df, 
    method="lm",
    trControl = trainControl(method = "cv", number = 10),
    preProcess = c("center", "scale"),
    metric = "RMSE"
  )
  return(ols_model)
}

get_ols_coefficient<- function(ols_model){
  b_ols<-coef(ols_model$finalModel)
  b_ols<- as.data.frame(as.matrix(b_ols))%>%
    mutate(Model= "OLS",Coefficient = row_number())
  return(b_ols)
}
```

- `RIDGE`

```{r, eval=FALSE}
# RIDGE
get_ridge_model<- function(df, lambda_grid){
  set.seed(45)
  ridge_model<-train(
    y~.-1,
    data = df,
    method = "glmnet",
    trControl = trainControl(method = "cv", number = 10, selectionFunction = "best"),
    preProcess = c("center", "scale"),
    tuneGrid = data.frame(alpha = 0, lambda = lambda_grid),
    metric = "RMSE"
  )
  return(ridge_model)
}

get_ridge_coefficient<- function(ridge_model){
  b_ridge<- coef(ridge_model$finalModel, ridge_model$bestTune$lambda)
  b_ridge<- as.data.frame(as.matrix(b_ridge))%>%
    mutate(Model = "RIDGE",Coefficient = row_number())
  return(b_ridge)
}

get_ridge_tune<- function(ridge_model){
  ridge_tune<-ridge_model$bestTune%>%
    mutate(Model = "RIDGE")
  return(ridge_tune)
}
```

- `LASSO`

```{r, eval=FALSE}
#LASSO
get_lasso_model<- function(df, lambda_grid){
  set.seed(45)
  lasso_model <- train(
    y~.-1,
    data = df,
    method = "glmnet",
    preProcess = c("center","scale"),
    trControl = trainControl(method = "cv", number = 10, selectionFunction = "best"),
    tuneGrid = data.frame(alpha = 1, lambda = lambda_grid),
    metric = "RMSE"
  )
  return(lasso_model)
}

get_lasso_coefficient<- function(lasso_model){
  b_lasso<- coef(lasso_model$finalModel, lasso_model$bestTune$lambda)
  b_lasso<- as.data.frame(as.matrix(b_lasso))%>%
    mutate(Model = "LASSO",Coefficient = row_number())
  return(b_lasso)
}

get_lasso_tune<- function(lasso_model){
  lasso_tune<-lasso_model$bestTune%>%
    mutate(Model = "LASSO")
  return(lasso_tune)
}
```

- `ELASTIC NET`


```{r, eval=FALSE}
#ELASTIC NET
get_elastic_net_model <- function(df, srchGrd){
  set.seed(45)
  elastic_net_model<-train(
    y~.-1,
    data = df,
    method = "glmnet",
    trControl = trainControl(method = "cv", number = 10, selectionFunction = "best"),
    tuneGrid = srchGrd,
    preProcess = c("center","scale"),
    metric = "RMSE"
  )
  return(elastic_net_model)
}

get_elastic_net_coefficient<- function(elastic_net_model){
  b_elastic_net<- coef(elastic_net_model$finalModel, elastic_net_model$bestTune$lambda)
  b_elastic_net<- as.data.frame(as.matrix(b_elastic_net))%>%
    mutate(Model = "ELASTIC NET", Coefficient = row_number())
  return(b_elastic_net)
}

get_elastic_net_tune<- function(elastic_net_model){
  elastic_net_tune<-elastic_net_model$bestTune%>%
    mutate(Model = "ELASTIC NET")
  return(elastic_net_tune)
}
```


## TRYOUT THE SIMULATION FOR ONE ITERATION

```{r, eval=FALSE}
#one simulation tryout
#inputs in the simulation,betas, b_final

df<-simulation(2,betas)

lambda_grid<- lambda_grid<-10^seq(-3, 3, length.out = 50)
alpha_grid <- c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1)
srchGrd = expand.grid(.alpha = alpha_grid, .lambda = lambda_grid)

ols_model<- get_ols_model(df)
ridge_model<- get_ridge_model(df, lambda_grid)
lasso_model<- get_lasso_model(df, lambda_grid)
elastic_net_model<- get_elastic_net_model(df, srchGrd)

b_ols<-get_ols_coefficient(ols_model)
colnames(b_ols)[1]<-1
b_ridge<-get_ridge_coefficient(ridge_model)
b_lasso<-get_lasso_coefficient(lasso_model)
b_elastic_net<- get_elastic_net_coefficient(elastic_net_model)

coefficients_all<-rbind(b_ols,b_ridge,b_lasso,b_elastic_net)

coefficient_column_name<-paste("simulation_",i, sep = "")

colnames(coefficients_all)[1]<- coefficient_column_name

b_final_tryout<-b_final%>%
  left_join(coefficients_all)
```

## SIMULATION FOR MULTIPLE ITERATIONS

```{r, eval=FALSE}
simulation_final_all_betas<-function(iterations, betas, b_final){
  lambda_grid<- lambda_grid<-10^seq(-3, 3, length.out = 50)
  alpha_grid <- c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1)
  srchGrd = expand.grid(.alpha = alpha_grid, .lambda = lambda_grid)
  for (i in seq(1, iterations, 1)){
    df<-simulation(i,betas)
    
    ols_model<- get_ols_model(df)
    ridge_model<- get_ridge_model(df, lambda_grid)
    lasso_model<- get_lasso_model(df, lambda_grid)
    elastic_net_model<- get_elastic_net_model(df, srchGrd)
    
    b_ols<-get_ols_coefficient(ols_model)
    colnames(b_ols)[1]<-1
    b_ridge<-get_ridge_coefficient(ridge_model)
    b_lasso<-get_lasso_coefficient(lasso_model)
    b_elastic_net<- get_elastic_net_coefficient(elastic_net_model)
      
    b_all<-rbind(b_ols,b_ridge,b_lasso,b_elastic_net)
    
    coefficient_column_name<-paste("simulation_",i, sep = "")
    colnames(b_all)[1]<- coefficient_column_name
    
    b_final<-b_final%>%
      left_join(b_all)
    }
  
  return(b_final)
}
  
```


For this study we decided to use 100 iterations, if we had more time we could have used more iterations. You could change the `100` to any number of iterations you would like, and the simulation will output all the coefficient values for 100 iterations for each of the 4 models. 

```{r, eval=FALSE}
simulation_all<-simulation_final_all_betas(100, betas, b_final)
```

```{r, eval=FALSE}
#create dataset for betas
write.csv(simulation_all, "simulation_all.csv")
```

# TUNING PARAMETERS


```{r}
tuning_final<-data_frame(matrix(ncol = 3, nrow = 0))

colnames(tuning_final)<-c("alpha","lambda","Model")

tuning_final
```


```{r}
simulation_final_tuning_parameters<-function(iterations, betas, tuning_final){
  for (i in seq(1, iterations, 1)){
    lambda_grid<- lambda_grid<-10^seq(-3, 3, length.out = 50)
    alpha_grid <- c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1)
    srchGrd = expand.grid(.alpha = alpha_grid, .lambda = lambda_grid)
    
    df<-simulation(seed = i,betas)
    
    ridge_model<- get_ridge_model(df, lambda_grid)
    lasso_model<- get_lasso_model(df, lambda_grid)
    elastic_net_model<- get_elastic_net_model(df, srchGrd)

    ridge_tune<-get_ridge_tune(ridge_model)
    lasso_tune<-get_lasso_tune(lasso_model)
    elastic_net_tune<- get_elastic_net_tune(elastic_net_model)
    
    head(ridge_tune)
    head(lasso_tune)
    head(elastic_net_tune)
    
    tune_all<-rbind(ridge_tune,lasso_tune,elastic_net_tune)%>%
      mutate(Simulation_number = i)
    
    tuning_final<- rbind(tuning_final, tune_all)
  }
  
  return(tuning_final)
}

```

```{r}
tuning_all<-simulation_final_tuning_parameters(100, betas, tuning_final)
```

```{r}
write.csv(tuning_all, "tuning_all.csv")
```

